Learning to Synthesize Rules for Network Intrusion Detectors

### 2020-02-25 paper/paper.tex
pass on eval (stop at 5.1)

\textbf{Implementation details.} We implemented function
\emph{isPlausible} by 1)~updating Suricata to only monitor the rule
that is passed as parameter, 2)~replaying the negative traffic (see
Section~\ref{sec:dataset-benign}), and 3)~reading the Suricata logs to
check if any benign message was captured by the rule. It is worth
noting that the benign traffic is stored in ``pcap'' files~\cite{pcap}
and Suricata can process those files very efficiently compared to
analyzing network messages directly. Instead of sending traffic to a
machine with a deployed server (\eg{}, web server) that
\suri\ monitors, our setup calls Suricata directly through an
API. Suricata is invoked multiple times in the process of minimizing
the rule.
\textbf{Limitations.} It is worth noting that although
\tname{} is accurate for generating rules that discrimate positive and
negative traffic, the quality of these rules is limited by the quality
of the training data it uses. For the positive traffic, \tname{} only
requires one sample, but can benefit of several samples. The same
applies for the negative traffic. In this case, however, \tname{} can
use public data sets available online.

\subsection{Step 3: Ranking}

@@

\newcommand{\textRQone}{How well \tname\ ranks the golden rule?}
\vspace{0.2cm}
\begin{itemize}[leftmargin=*]
\item{\textbf{RQ1.}} \textRQone
\end{itemize}

\noindent
\textbf{Rationale.}~The golden rule is a trusted human-written rule
obtained from credible sources. \tname\ generates several rules that
satisfy the invariant to only capture the malicious traffic, but the
amount of data we use to evaluate that property, albeit large, is
limited. Assessing the quality of each rule \tname\ produces (\ie{},
fitness to the attack) requires domain knowledge and much effort. The
position of the golden rule at the ranking produced by \tname\ is a
proxy measure to the effort a system administrator would have in
finding the rule that fits the purpose.

\newcommand{\textRQtwo}{What is the importance of heuristic functions in ranking?}
\vspace{0.2cm}
\begin{itemize}[leftmargin=*]
\item{\textbf{RQ2.}} \textRQtwo\
\end{itemize}

\noindent
\textbf{Rationale.}~\tname\ proposes a variety of heuristic functions
to rank the plausible rules produced during minimization. It is
therefore important to understand how much each function contributes
to the overall performance of the technique.

\newcommand{\textRQthree}{What is the importance of the positive and
  negative traffic in ranking?}
\vspace{0.2cm}
\begin{itemize}[leftmargin=*]
\item{\textbf{RQ3.}} \textRQthree\
\end{itemize}

\noindent
\textbf{Rationale.}~\tname\ uses both positive and negative examples
to generate rules. This question investigates how sensible the
technique is to the traffic that it uses.
\noindent
\vspace{1ex}

The following sections describe the objects we used to evaluate the
technique (Section~\ref{sec:dataset-benign}), answer the questions
(sections~\ref{sec:answer-rqone},~\ref{sec:answer-rqothree}), and
discuss threats to the validity of the
experiments~(Section~\ref{sec:threats}).

@@

\subsubsection{Malicious traffic.}To reproduce the malicious traffic, we
configured an HTTP server (Apache2) in one machine and ran
Nikto2~\cite{nikto} in a different machine. We used this setup to
replay each of the attacks from Table~\ref{table:attacks}.  On each
execution, we captured the incoming traffic to the web server with
Wireshark~\cite{wireshark-net-monitor} and used the corresponding pcap
files as an input to \tname.\Mar{@Lucas, eu traduzi o que vc. falou
  para ingles. Mas, aparentemente, isto esta **inconsistente** com o
  que eh dito na Secao~\ref{attack-reproduction}. Por favor, esclareca!}
  \Luc{Essa descrição é referente aos ataques HTTP/TCP da tabela. O
    hping3 gera apenas o trafego do SYN Flood. Ambas as explicações
    estão corretas. Creio q a descrição da seção 5.2.1 tenha q ser
    movida para cá pois está explicando como capturamos o trafego
    malicioso. Nós executamos o Nikto e o hping3 uma unica vez para
    capturar com o wireshark o trafego q eles geram. O trafego
    produzido pelo Nikto ja inclui todos os ataques que utilizamos,
    nós separamos os pacotes de cada um deles manualmente utilizando o
    wireshark.}\Mar{Fiquei na duvida se o hping soh serve para o SYN
    Flood e se/porque o Nikto nao eh usado neste caso...Enfim, nao
    precisa responder por aqui. Mova o texto deixando-o
    *consistente*.}

  \tname{} requires the attack to
be replayed in order to check whether or not the rule it generates
captures the attack.  See function $\mathit{isPlausible}$, used in the
minimization algorithm from Section~\ref{sec:minimization}. For all
attacks listed on Table~\ref{table:attacks}, we found scripts
available on the web to replay the attacks. For example, the SYN Flood
attack can be replayed with hping3~\cite{hping3}, using the following
command: \CodeIn{hping3 -d 80 -w 64 -S -p 80 --flood --rand-source
  <IP>}. With the script to reproduce the attack, we executed the
following steps to obtain ``pcap'' files, which abstracts network
messages~\cite{pcap}, to run \tname:~1)~instantiate the script (\eg{},
define the IP, for the example above), 2)~execute the script,
3)~monitor the traffic with the wireshark network
monitor~\cite{wireshark-net-monitor}.

@@

\subsection{Answering RQ2: \textRQtwo}
\label{sec:answer-rqotwo}

\subsection{Answering RQ3: \textRQthree}
\label{sec:answer-rqothree}

>>>>

\textbf{Implementation details.} We implemented function
\emph{isPlausible} by 1)~updating Suricata to only monitor the rule
that is passed as parameter, 2)~replaying the negative traffic (see
Section~\ref{sec:dataset-benign}), and 3)~reading the Suricata logs to
check if any benign message was captured by the rule. It is worth
noting that the benign traffic is stored in \emph{pcap}
files~\cite{pcap} and Suricata can process those files very
efficiently compared to analyzing network messages directly. Instead
of sending traffic to a machine with a deployed server (\eg{}, web
server) that \suri\ monitors, our setup calls Suricata directly
through an API. Suricata is invoked multiple times in the process of
minimizing the rule.  \textbf{Limitations.} It is worth noting that
although \tname{} is accurate for generating rules that discrimate
positive and negative traffic, the quality of these rules is limited
by the quality of the training data it uses. For the positive traffic,
\tname{} only requires one sample, but can benefit of several
samples. The same applies for the negative traffic. In this case,
however, \tname{} can use public data sets available online.

\subsection{Step 3: Ranking}
\label{sec:ranking}

@@

\newcommand{\textRQone}{How well \tname\ ranks the golden rule?}
\vspace{0.2cm}
\begin{itemize}[leftmargin=*,label={}]
\item{\textbf{RQ1.}} \textRQone
\end{itemize}

\noindent
\textbf{Rationale.}~The golden rule is a trusted human-written
\nids\ rule that we obtained from credible sources. \tname\ generates
several rules that satisfy the invariant to only capture the malicious
traffic, but the amount of data we used to evaluate that property,
albeit large, is limited. Assessing the quality of each rule
\tname\ produces (\ie{}, fitness to the attack) requires domain
knowledge and much effort. The position of the golden rule at the
ranking produced by \tname\ is a proxy measure to the effort of a
system administrator in finding the rule that fits the purpose.

\newcommand{\textRQtwo}{What is the importance of heuristic functions
  on \tname's performance?}
\vspace{0.2cm}
\begin{itemize}[leftmargin=*,label={}]
\item{\textbf{RQ2.}} \textRQtwo\
\end{itemize}

\noindent
\textbf{Rationale.}~We proposed a variety of heuristic functions to
rank the plausible rules that \tname\ created during minimization (as
per step 2 in the workflow). Recall that these functions are based on
the data from a public ruleset~\cite{emerging-threats-open} (see
Section~\ref{sec:ranking}). It is therefore important to understand
the extent to which each function contributes to the overall
performance of the technique.

\newcommand{\textRQthree}{What is the importance of the benign and
  malicious traffic on \tname's performance?}
\vspace{0.2cm}
\begin{itemize}[leftmargin=*,label={}]
\item{\textbf{RQ3.}} \textRQthree\
\end{itemize}

\noindent
\textbf{Rationale.}~\tname\ uses both positive and negative examples
to generate rules. This question investigates how sensible the
technique is to the selection of these inputs.
\noindent
\vspace{1ex}

The following sections describe the objects we used to evaluate the
technique (Section~\ref{sec:dataset-benign}), answer the posed
research questions (sections~\ref{sec:answer-rqone},
and~\ref{sec:answer-rqtwo}, and~\ref{sec:answer-rqthree}), and
discuss threats to the validity of the
experiments~(Section~\ref{sec:threats}).

@@

\subsubsection{Malicious traffic.}To reproduce the malicious traffic
for http and tcp attacks, we configured an http server (Apache2) in
one machine and Nikto2~\cite{nikto} in a different machine. Nikto is
an open source tool with a catalog of scripts to reproduce
vulnerabilities. It is important to note that this setup is \emph{not}
necessary if \emph{pcap} files---pcap is the input format used by
\tname{}---for the attack already exist. As explained in
Section~\ref{sec:minimization}, pcap files abstract network
messages~\cite{pcap} and can be used to avoid network traffic during
rule synthesis. We proceeded as follows to obtain pcap files for the
the attacks we analyzed: we 1)~instantiated the script (\eg{}, defined
the IP of the web server), 2)~executed the Nikto2 script, 3)~and
monitored the traffic with the Wireshark network
monitor~\cite{wireshark-net-monitor}. The Wireshark tool records the
traffic and reports pcap files on output for replaying. The only
difference in the setup of non-http non-tcp attack is in the first
step of this procedure. For example, we used the hping3~\cite{hping3}
command to reproduce the ICMP SYN Flood attack: \CodeIn{hping3 -d 80
  -w 64 -S -p 80 --flood --rand-source <IP>}. This command alone
reproduces a flooding attack on a given machine
(Section~\ref{sec:dos}).\Mar{what about the variants? where do they
  come from? this is confusing as the narrative above suggests the
  attack is hardcoded in the tool.}

%% On each execution, we captured
%% the incoming traffic to the web server with
%% Wireshark~\cite{wireshark-net-monitor} and used the corresponding pcap
%% files as an input to

%% \tname.\Mar{@Lucas, eu traduzi o que vc. falou
%%   para ingles. Mas, aparentemente, isto esta **inconsistente** com o
%%   que eh dito na Secao~\ref{attack-reproduction}. Por favor,
%%   esclareca!}  \Luc{Essa descrição é referente aos ataques HTTP/TCP da
%%   tabela. O hping3 gera apenas o trafego do SYN Flood. Ambas as
%%   explicações estão corretas. Creio q a descrição da seção 5.2.1 tenha
%%   q ser movida para cá pois está explicando como capturamos o trafego
%%   malicioso. Nós executamos o Nikto e o hping3 uma unica vez para
%%   capturar com o wireshark o trafego q eles geram. O trafego produzido
%%   pelo Nikto ja inclui todos os ataques que utilizamos, nós separamos
%%   os pacotes de cada um deles manualmente utilizando o
%%   wireshark.}\Mar{Fiquei na duvida se o hping soh serve para o SYN
%%   Flood e se/porque o Nikto nao eh usado neste caso...Enfim, nao
%%   precisa responder por aqui. Mova o texto deixando-o *consistente*.}
%%   \tname{} requires the attack to
%% be replayed in order to check whether or not the rule it generates
%% captures the attack.  See function $\mathit{isPlausible}$, used in the
%% minimization algorithm from Section~\ref{sec:minimization}. For all
%% attacks listed on Table~\ref{table:attacks}, we found scripts
%available on the web to replay the attacks. For example,

\Mar{--------------------------- estou aqui!}

@@

\subsection{Answering RQ2: \textRQtwo}
\label{sec:answer-rqtwo}

\subsection{Answering RQ3: \textRQthree}
\label{sec:answer-rqthree}
